---
title: "Topic 3_non synchronized A549 mit peak finder"
author: "Johannes Schadt, Anastasia Möller, Sylviane Verschaeve, Tine Limberg"
date: "2023-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Proteome-wide Screen for RNA-dependent Proteins:

## *non-synchronized A549 cells*

Zum PDF exportieren:

```{r}
#tinytex::install_tinytex()
```

Loading the data:

```{r}
#MS_Table <- read.csv('RDeep_A549_NS.csv', header=TRUE, row.names=1, sep = ",")
#View(MS_Table)
MS_Table <- read.delim('https://www.dropbox.com/s/vm3lxljjm9chau8/RDeeP_A549_NS.csv?dl=1', header=TRUE, row.names=1, sep = ";")
```

## 1. Preparing data for analysis

#### 0.1 Data description

```{r}
dim(MS_Table)
head(rownames(MS_Table))
head(colnames(MS_Table))
```

#### 1.1 Check for missing values

```{r}
sum(apply(MS_Table, 1, anyNA)) == 0
sum(is.na(MS_Table)) == 0
```

#### 1.2 Check data format

```{r}
MS_Table = apply(MS_Table,2,as.numeric)
sum(apply(MS_Table, 1, is.numeric)) == nrow(MS_Table)
```

#### 1.3 Deleting rows with only zeros

```{r}
min(MS_Table)
sum(apply(MS_Table,1,sum)==0)
```

-\> da die Summe der Zeileneinträge keines Proteins 0 entspricht, wurde ein Dataframe aus False erstellt. Einträge ausschließlich False, werden durch die sum Funktion als 0 aufaddiert.

#### 1.4.1 Reordering columns

```{r}
MS_Table_reordered <- MS_Table[, c(
  paste0("Fraction", 1:25, "_Ctrl_Rep1"),
  paste0("Fraction", 1:25, "_Ctrl_Rep2"),
  paste0("Fraction", 1:25, "_Ctrl_Rep3"),
  paste0("Fraction", 1:25, "_RNase_Rep1"),
  paste0("Fraction", 1:25, "_RNase_Rep2"),
  paste0("Fraction", 1:25, "_RNase_Rep3")
)]

#View(MS_Table_reordered)
sum(apply(MS_Table_reordered, 2, is.numeric)) == ncol(MS_Table)
```

#### 1.4.2 Separate Ctrl and RNase

```{r}
MS_Table_Ctrl <-MS_Table_reordered[,1:75]
#View(MS_Table_Ctrl)
MS_Table_RNase <-MS_Table_reordered[,76:150]
#View(MS_Table_RNase)
```

## 2. Normalize, Reduce and Scale Dataset

#### 2.1 Normalization (z-Transformation)

For the normalization each replicate has to be separated, therefore we design 6 separate dataframes.

```{r}
MS_Table_Ctrl_rep1 <-MS_Table_reordered[,1:25]
MS_Table_Ctrl_rep2 <-MS_Table_reordered[,26:50]
MS_Table_Ctrl_rep3 <-MS_Table_reordered[,51:75]
MS_Table_RNase_rep1 <-MS_Table_reordered[,76:100]
MS_Table_RNase_rep2 <-MS_Table_reordered[,101:125]
MS_Table_RNase_rep3 <-MS_Table_reordered[,126:150]
```

Normalization with z-Transformation: We use the scale command. EXPLANATION WHY Z-TRANSFORMATION !!

NOTE to better understand the sweep function.

```{r}
m <- matrix(1:6, nrow = 2)
m
# Subtract the mean of each column from the corresponding elements in the matrix
swept_matrix <- sweep(m, 2, colMeans(m), FUN = "-")

print(swept_matrix)

applied_matrix <- apply(m, 2, function(x) x - mean(x))

print(applied_matrix)
```

First the normalization for the Ctrl: 

```{r}
ctrl1_meanvalue = t(apply(MS_Table_Ctrl_rep1,1,function(x){x-mean(x)}))
View(ctrl1_meanvalue)

```

```{r}
ctrl_z_transformed = t(scale(t(MS_Table_Ctrl)))

ctrl1_z_transformed = ctrl_z_transformed[,1:25]
ctrl2_z_transformed = ctrl_z_transformed[,26:50]
ctrl3_z_transformed = ctrl_z_transformed[,51:75]

#calculate the min value for each protein
min_ctrl1_z_transformed = apply(ctrl1_z_transformed, 1, min)
min_ctrl2_z_transformed = apply(ctrl2_z_transformed, 1, min)
min_ctrl3_z_transformed = apply(ctrl3_z_transformed, 1, min)

ctrl1_pos.= sweep(ctrl1_z_transformed,1,min_ctrl1_z_transformed,FUN = '-')
ctrl2_pos.= sweep(ctrl2_z_transformed,1,min_ctrl2_z_transformed,FUN = '-')
ctrl3_pos.= sweep(ctrl3_z_transformed,1,min_ctrl3_z_transformed,FUN = '-')

View(ctrl1_pos.)
View(ctrl1_pos.)
View(ctrl1_pos.)
```

Still with the z-transformation we took for each protein the same mean over all values. Since the protein amount in each replicate is different, it is better to calculate the mean extra for each replicate. However the sd-value has not to be adapted. Because the replicates have the same variance (same procedure for every replicate in the wet lab), we don't have to calculate the standarddevidation for the replicates in each fraction extra. We can calculate the sd-value for one row /sd.

*sd-values and mean values of Ctrl:*

```{r}
sd_MS_Table_Ctrl = apply(MS_Table_Ctrl, 1, sd)

mean_MS_Table_Ctrl_rep1 = apply(MS_Table_Ctrl_rep1, 1, mean)
mean_MS_Table_Ctrl_rep2 = apply(MS_Table_Ctrl_rep2, 1, mean)
mean_MS_Table_Ctrl_rep3 = apply(MS_Table_Ctrl_rep3, 1, mean)
```

*Normalization of Ctrl:*

```{r}
ctrl1_meanvalue = sweep(MS_Table_Ctrl_rep1,1,mean_MS_Table_Ctrl_rep1,'-')
ctrl1_normalized = sweep(ctrl1_meanvalue,1,sd_MS_Table_Ctrl,'/')

ctrl2_meanvalue = sweep(MS_Table_Ctrl_rep2,1,mean_MS_Table_Ctrl_rep2,'-')
ctrl2_normalized = sweep(ctrl2_meanvalue,1,sd_MS_Table_Ctrl,'/')

ctrl3_meanvalue = sweep(MS_Table_Ctrl_rep3,1,mean_MS_Table_Ctrl_rep3,'-')
ctrl3_normalized = sweep(ctrl3_meanvalue,1,sd_MS_Table_Ctrl,'/')

#calculate again the min value for each protein
min_ctrl1_normalized = apply(ctrl1_normalized, 1, min)
min_ctrl2_normalized = apply(ctrl2_normalized, 1, min)
min_ctrl3_normalized = apply(ctrl3_normalized, 1, min)

#substract the min value from each position to discard the 0
ctrl1_normalized_pos.= 
  sweep(ctrl1_normalized,1,min_ctrl1_normalized,FUN = '-')
ctrl2_normalized_pos.= 
  sweep(ctrl2_normalized,1,min_ctrl2_normalized,FUN = '-')
ctrl3_normalized_pos.= 
  sweep(ctrl3_normalized,1,min_ctrl3_normalized,FUN = '-')
```

*sd-values and mean values of RNase:*

```{r}
sd_MS_Table_RNase = apply(MS_Table_RNase, 1, sd)

mean_MS_Table_RNase_rep1 = apply(MS_Table_RNase_rep1, 1, mean)
mean_MS_Table_RNase_rep2 = apply(MS_Table_RNase_rep2, 1, mean)
mean_MS_Table_RNase_rep3 = apply(MS_Table_RNase_rep3, 1, mean)
```

*Normalization of Ctrl:*

```{r}
rnase1_meanvalue = sweep(MS_Table_RNase_rep1,1,mean_MS_Table_RNase_rep1,'-')
rnase1_normalized = sweep(rnase1_meanvalue,1,sd_MS_Table_RNase,'/')

rnase2_meanvalue = sweep(MS_Table_RNase_rep2,1,mean_MS_Table_RNase_rep2,'-')
rnase2_normalized = sweep(rnase2_meanvalue,1,sd_MS_Table_RNase,'/')

rnase3_meanvalue = sweep(MS_Table_RNase_rep3,1,mean_MS_Table_RNase_rep3,'-')
rnase3_normalized = sweep(rnase3_meanvalue,1,sd_MS_Table_RNase,'/')

#calculate again the min value for each protein
min_rnase1_normalized = apply(rnase1_normalized, 1, min)
min_rnase2_normalized = apply(rnase2_normalized, 1, min)
min_rnase3_normalized = apply(rnase3_normalized, 1, min)

#substract the min value from each position to discard the 0
rnase1_normalized_pos.= 
  sweep(rnase1_normalized,1,min_rnase1_normalized,FUN = '-')
rnase2_normalized_pos.= 
  sweep(rnase2_normalized,1,min_rnase2_normalized,FUN = '-')
rnase3_normalized_pos.= 
  sweep(rnase3_normalized,1,min_rnase3_normalized,FUN = '-')
```

#### 2.2 Reduce

To reduce we take the mean value between each replicate.

```{r}
ctrl_reduced = 
  (ctrl1_normalized_pos. + ctrl2_normalized_pos. + ctrl3_normalized_pos.)/3
rnase_reduced = 
  (rnase1_normalized_pos. + rnase2_normalized_pos. + rnase3_normalized_pos.)/3
```

#### 2.3 Scaling to 100

*Each Replicate of Ctrl:*

```{r}
ctrl1_scaled = 
  sweep(ctrl1_normalized_pos.,1,100/rowSums(ctrl1_normalized_pos.),'*')
# rowSums(ctrl1_scaled)
ctrl2_scaled = 
  sweep(ctrl2_normalized_pos.,1,100/rowSums(ctrl2_normalized_pos.),'*')
# rowSums(ctrl2_scaled)
ctrl3_scaled = 
  sweep(ctrl3_normalized_pos.,1,100/rowSums(ctrl3_normalized_pos.),'*')
# rowSums(ctrl3_scaled)
```

Reduced Ctrl:

```{r}
ctrl_scaled = 
  sweep(ctrl_reduced,1,100/rowSums(ctrl_reduced),'*')
# rowSums(ctrl_scaled)
```

*Each Replicate of RNase:*

```{r}
rnase1_scaled = 
  sweep(rnase1_normalized_pos.,1,100/rowSums(rnase1_normalized_pos.),'*')
# rowSums(rnase1_scaled)
rnase2_scaled = 
  sweep(rnase2_normalized_pos.,1,100/rowSums(rnase2_normalized_pos.),'*')
# rowSums(rnase2_scaled)
rnase3_scaled = 
  sweep(rnase3_normalized_pos.,1,100/rowSums(rnase3_normalized_pos.),'*')
# rowSums(rnase3_scaled)
```

*Reduced RNase*

```{r}
rnase_scaled = 
  sweep(rnase_reduced,1,100/rowSums(rnase_reduced),'*')
# rowSums(rnase_scaled)
```

##3. Peakfinder

```{sd_by_row <- apply(rnase.mms, 1, function(x) sd(rnase.mms[x < 20]))}
mean(sd_by_row)





{r}
rnase3_peaks <- rnase3_scaled

colnames(rnase3_peaks) <- paste0("Rep_3_RnaseFraction", seq_along(colnames(rnase3_peaks)) - 1)
maxpeaks <- apply(rnase3_scaled, 1, function(x){which.max(x)
  })

peaks <- apply(rnase3_scaled, 1,
               function(x) {
                 peak_indices <- which(diff(sign(diff(x)))==-2 & x[-c(1,length(x))] > 3) + 1
                 peak_sd <- sapply(peak_indices,
                                    function(i) sd(x[(i-1):(i+1)]))
                 peak_indices[peak_sd > 0.2488523] #wenn standardabweichung kleiner als 1 wird Wert nicht als Maximum angesehen
                
               

               })
View(peaks)

largevalues <-apply(rnase3_scaled, 1,function(x) {largevalues_indices <- which(x > 9)
                 peak_mean <- sapply(largevalues_indices,
                                    function(i) mean(rnase3_scaled[(i-1):(i+1)]))
                 largevalues_indices[peak_mean >9] 
               unique(c(largevalues_indices - 1, largevalues_indices, largevalues_indices + 1))
})


localmaxima <- peaks
for (i in seq_along(localmaxima)) {
    localmaxima[[i]] <- localmaxima[[i]][!localmaxima[[i]] %in% maxpeaks[[i]]]
}
View(maxpeaks)
View(localmaxima)
View(rnase3_scaled)
rnase3_scaled <- as.data.frame(rnase3_scaled)
#View(largevalues)
#Randwerte überprüfen 
if (ncol(rnase3_scaled) >= 2) {
  result <- lapply(as.data.frame(t(rnase3_scaled)), function(x) {
    if (!any(is.na(x)) & (x[1] > x[2] | x[nrow(rnase3_scaled)] > x[nrow(rnase3_scaled) - 1])) {
      return(c(1, nrow(rnase3_scaled)))
    } else if (!any(is.na(x)) & x[1] > x[2]) {
      return(1)
    } else if (!any(is.na(x)) & x[nrow(rnase3_scaled)] > x[nrow(rnase3_scaled) - 1]) {
      return(nrow(rnase3_scaled))
    } else {
      return(NULL)
    }
  })
} else {
}

neighbors <- apply(rnase3_scaled, 1,
               function(x) {
                 peak_indices <- which(diff(sign(diff(x))) == -2 & x[-c(1, length(x))] > 3) + 1
                 peak_sd <- sapply(peak_indices,
                                    function(i) sd(x[(i-1):(i+1)]))
                 peak_indices <- peak_indices[peak_sd > 0.2488523]
                 c(peak_indices - 1, peak_indices, peak_indices + 1)
})
View(neighbors)
#Probleme Bennung der Spalten um eins verschoben (keine Ahnung warum???)
#Zeile 786 erkennt kein Maxima, da Spalte 25 größer als 24 muss noch behoben werden
#

```

```{r}
shoulderregion <- largevalues                  
for (i in seq_along(shoulderregion)) {
    shoulderregion[[i]] <- shoulderregion[[i]][!shoulderregion[[i]] %in% neighbors[[i]]]
}
View(largevalues)
View(shoulderregion)
View(neighbors)
```
