```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```

## Proteome-wide Screen for RNA-dependent Proteins:

## *non-synchronized A549 cells*

Zum PDF exportieren

```{r}
# tinytex::install_tinytex()
```

Loading the data:

```{r}
MS_Table <- read.delim('https://www.dropbox.com/s/vm3lxljjm9chau8/RDeeP_A549_NS.csv?dl=1', header=TRUE, row.names=1, sep = ";")
```

## 1. Preparing data for analysis

#### 0.1 Data description

```{r}
dim(MS_Table)
head(rownames(MS_Table))
head(colnames(MS_Table))
```

#### 1.1 Check for missing values

```{r}
sum(apply(MS_Table, 1, anyNA)) == 0
sum(is.na(MS_Table)) == 0
```

#### 1.2 Check data format

```{r}
sum(apply(MS_Table, 1, is.numeric)) == nrow(MS_Table)
```

#### 1.3 Deleting rows with only zeros

```{r}
min(MS_Table)
sum(apply(MS_Table,1,sum)==0)
```

-\> da die Summe der Zeileneinträge keines Proteins 0 entspricht, wurde ein Dataframe aus False erstellt. Einträge ausschließlich False, werden durch die sum Funktion als 0 aufaddiert.

#### 1.4.1 Reordering columns

```{r}
MS_Table_reordered <- MS_Table[, c(
  paste0("Fraction", 1:25, "_Ctrl_Rep1"),
  paste0("Fraction", 1:25, "_Ctrl_Rep2"),
  paste0("Fraction", 1:25, "_Ctrl_Rep3"),
  paste0("Fraction", 1:25, "_RNase_Rep1"),
  paste0("Fraction", 1:25, "_RNase_Rep2"),
  paste0("Fraction", 1:25, "_RNase_Rep3")
)]
# View(MS_Table_reordered)
sum(apply(MS_Table_reordered, 2, is.numeric)) == ncol(MS_Table)
```

#### 1.4.2 Separate Ctrl and RNase

```{r}
MS_Table_Ctrl <-MS_Table_reordered[,1:75]
#View(MS_Table_Ctrl)
MS_Table_RNase <-MS_Table_reordered[,76:150]
#View(MS_Table_RNase)
```

## 2. Scale and Reduce Dataset

#### 2.1 Normalization (mean value methode)

For the normalization each replicate has to be separated, therefore we design 6 separate dataframes.

```{r}
MS_Table_Ctrl_rep1 <-MS_Table_reordered[,1:25]
MS_Table_Ctrl_rep2 <-MS_Table_reordered[,26:50]
MS_Table_Ctrl_rep3 <-MS_Table_reordered[,51:75]
MS_Table_RNase_rep1 <-MS_Table_reordered[,76:100]
MS_Table_RNase_rep2 <-MS_Table_reordered[,101:125]
MS_Table_RNase_rep3 <-MS_Table_reordered[,126:150]
```

On them we apply separately the mean value methode.

```{r}
#rowMeans(MS_Table_Ctrl_rep1)
#apply(MS_Table_Ctrl_rep1,1,sd)
MS_Table_Ctrl_rep1_normalized <- apply(t(MS_Table_Ctrl_rep1),1, function(x)(x - mean(x)) / sd(x))
View(MS_Table_Ctrl_rep1_normalized)
# Der oben stehende code führt nicht die mean value methode durch. Daher siehe unten.  
```

Another idea to apply the mean value method on each row of every dataframe. All values below 0 are set back to 0. 
```{r}

# MS_Table_Ctrl_rep1
MS_Table_Ctrl_rep1_normalized <- t(apply(MS_Table_Ctrl_rep1, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  return(normalized)
}))

# View(MS_Table_Ctrl_rep1_normalized)

# MS_Table_Ctrl_rep2
MS_Table_Ctrl_rep2_normalized <- t(apply(MS_Table_Ctrl_rep2, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  return(normalized)
}))

# MS_Table_Ctrl_rep3
MS_Table_Ctrl_rep3_normalized <- t(apply(MS_Table_Ctrl_rep3, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  return(normalized)
}))

# MS_Table_RNase_rep1
MS_Table_RNase_rep1_normalized <- t(apply(MS_Table_RNase_rep1, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  return(normalized)
}))

# MS_Table_RNase_rep2
MS_Table_RNase_rep2_normalized <- t(apply(MS_Table_RNase_rep2, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  return(normalized)
}))

# MS_Table_RNase_rep3
MS_Table_RNase_rep3_normalized <- t(apply(MS_Table_RNase_rep3, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  return(normalized)
}))

```

Normalization with the mean value method and scaling to 100

```{r}
# MS_Table_Ctrl_rep1
MS_Table_Ctrl_rep1_normalized <- t(apply(MS_Table_Ctrl_rep1, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  scaled <- normalized * (100 / sum(normalized))
  return(scaled)
}))

# View(MS_Table_Ctrl_rep1_normalized)

# MS_Table_Ctrl_rep2
MS_Table_Ctrl_rep2_normalized <- t(apply(MS_Table_Ctrl_rep2, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  scaled <- normalized * (100 / sum(normalized))
  return(scaled)
}))

# MS_Table_Ctrl_rep3
MS_Table_Ctrl_rep3_normalized <- t(apply(MS_Table_Ctrl_rep3, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  scaled <- normalized * (100 / sum(normalized))
  return(scaled)
}))

# MS_Table_RNase_rep1
MS_Table_RNase_rep1_normalized <- t(apply(MS_Table_RNase_rep1, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  scaled <- normalized * (100 / sum(normalized))
  return(scaled)
}))

# MS_Table_RNase_rep2
MS_Table_RNase_rep2_normalized <- t(apply(MS_Table_RNase_rep2, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  scaled <- normalized * (100 / sum(normalized))
  return(scaled)
}))

# MS_Table_RNase_rep3
MS_Table_RNase_rep3_normalized <- t(apply(MS_Table_RNase_rep3, 1, function(x) {
  normalized <- x - mean(x)
  normalized[normalized < 0] <- 0
  scaled <- normalized * (100 / sum(normalized))
  return(scaled)
}))

```

Unite the replicates and scale back to 100 - Control 
```{r}
MS_Table_Ctrl_reduced = 
  (MS_Table_Ctrl_rep1_normalized + MS_Table_Ctrl_rep2_normalized + MS_Table_Ctrl_rep3_normalized)/3
MS_Table_Ctrl_scaled = 
  sweep(MS_Table_Ctrl_reduced,1,100/rowSums(MS_Table_Ctrl_reduced),'*')

# Check if the two data frames are identical
is_identical <- identical(MS_Table_Ctrl_reduced, MS_Table_Ctrl_scaled)

# Print the result
if (is_identical) {
  print("The data frames are identical.")
} else {
  print("The data frames are not identical.")
}
```
-> scaling back to 100 is necessary

Unite the replicates and scale back to 100 - RNase
```{r}

MS_Table_RNase_reduced = 
  (MS_Table_RNase_rep1_normalized + MS_Table_RNase_rep2_normalized + MS_Table_RNase_rep3_normalized)/3
MS_Table_RNase_scaled = 
  sweep(MS_Table_RNase_reduced,1,100/rowSums(MS_Table_RNase_reduced),'*')

```

creating line graphs 
```{r}
# Anzahl der Zeilen im Dataframe
num_rows <- nrow(MS_Table_RNase_scaled)

# Erstelle einen leeren Plot für jeden Graphen
plots <- list()

# Iteriere über jede Zeile und erstelle den stetigen Graphen
for (i in 1:num_rows) {
  row_data <- MS_Table_RNase_scaled[i, ]
  
  # Überprüfe, ob genügend nicht-NA-Werte vorhanden sind
  if (sum(!is.na(row_data)) >= 2) {
    x <- 1:length(row_data)
    y <- as.numeric(row_data)
    interp_data <- approx(x, y, method = "linear", n = 1000)
    plot(interp_data, type = "l", xlab = "X-Achse", ylab = "Y-Achse")
    
    # Speichere den Plot im List-Objekt
    plots[[i]] <- recordPlot()
    
    # Schließe das aktuelle Grafikgerät
    dev.off()
  } else {
    # Falls nicht genügend nicht-NA-Werte vorhanden sind, füge einen leeren Plot hinzu
    plots[[i]] <- NULL
  }
}

# Zugriff auf die einzelnen Plots
# Zum Beispiel: plots[[1]] für den ersten Plot, plots[[2]] für den zweiten Plot, usw.



```



An idea to check the reproducibility -> das hinzufügen der Funktion as.numeric hat das Problem gelöst
## Correlation Control rep 1&2
```{r}
 
num_rows <- nrow(MS_Table_Ctrl_rep1)

# create vector 
correlations <- numeric(num_rows) 


for (i in 1:num_rows) {
  # select row i from MS_Table_Ctrl_rep1 and MS_Table_Ctrl_rep2 
  row_rep1 <- as.numeric(MS_Table_Ctrl_rep1[i, ]) 
  # Der Fehler war, dass die Rows nicht als Vector, sondern als Dataframe eingelesen wurden -> Sobald der Type richtig ist, kann er auch die Correlation ausrechnen. 
  # Als weiterer Hinweis jedoch: Die Operation könnte etwas länger dauern, wenn ihrs für alle dfs macht. 
  row_rep2 <- as.numeric(MS_Table_Ctrl_rep2[i, ])
  
  
  correlation <- cor(row_rep1, row_rep2, method = "pearson")
  
  
  correlations[i] <- correlation
}

# show result 
correlations



```

shorter version to check reproducibility 
```{r}
ctrl_cor = 
  cbind(ctrl_cor_1_2 = 
          sapply(seq.int(dim(MS_Table_Ctrl_rep1)[1]), function(x) cor(as.numeric(MS_Table_Ctrl_rep1[x,]), as.numeric(MS_Table_Ctrl_rep2[x,]))),
        ctrl_cor_2_3 = 
          sapply(seq.int(dim(MS_Table_Ctrl_rep1)[1]), function(x) cor(as.numeric(MS_Table_Ctrl_rep2[x,]), as.numeric(MS_Table_Ctrl_rep3[x,]))), 
        ctrl_cor_1_3 = 
          sapply(seq.int(dim(MS_Table_Ctrl_rep1)[1]), function(x) cor(as.numeric(MS_Table_Ctrl_rep1[x,]), as.numeric(MS_Table_Ctrl_rep3[x,]))))

rnase_cor = 
  cbind(rnase_cor_1_2 = 
         sapply(seq.int(dim(MS_Table_RNase_rep1)[1]), function(x) cor(as.numeric(MS_Table_RNase_rep1[x,]), as.numeric(MS_Table_RNase_rep2[x,]))),
        ctrl_cor_2_3 = 
          sapply(seq.int(dim(MS_Table_Ctrl_rep1)[1]), function(x) cor(as.numeric(MS_Table_RNase_rep2[x,]), as.numeric(MS_Table_RNase_rep3[x,]))), 
        ctrl_cor_1_3 = 
          sapply(seq.int(dim(MS_Table_Ctrl_rep1)[1]), function(x) cor(as.numeric(MS_Table_RNase_rep1[x,]), as.numeric(MS_Table_RNase_rep3[x,]))))
```

```{r}
ctrl_na_rows <- which(rowSums(is.na(ctrl_cor)) > 0)
ctrl_na_rows

rnase_na_rows <- which(rowSums(is.na(rnase_cor)) > 0)
rnase_na_rows
length(rnase_na_rows)
```

In total we look at 3\*3680 correlation. This has to be taken into account, when looking at the graphs. It is import to figure out if the 3 cor are for one protein or for 3 different ones.

```{r}
library(ggplot2)
rnase_cor_data_frame = data.frame(c(rnase_cor[,1],rnase_cor[,2],rnase_cor[,3]))
colnames(rnase_cor_data_frame) = "correlation"
ggplot(rnase_cor_data_frame, aes(x=correlation)) + geom_histogram()
```

```{r}
ctrl_cor_data_frame = data.frame(c(ctrl_cor[,1],ctrl_cor[,2],ctrl_cor[,3]))
colnames(ctrl_cor_data_frame) = "correlation"
ggplot(ctrl_cor_data_frame, aes(x=correlation)) + geom_histogram()
```

```{r}
non_reproducible = unique(which(rowSums(ctrl_cor<0.95)>0), which(rowSums(rnase_cor<0.95)>0))

MS_Table_Ctrl_clean = MS_Table_Ctrl[-non_reproducible,]
MS_Table_RNase_clean = MS_Table_RNase[-non_reproducible,]

'View(MS_Table_Ctrl_clean)
```

