<<<<<<< HEAD
---
title: "Proteome-wide Screen for RNA-dependent Proteins: non-synchronized A549 cells"
author: "Johannes Schadt, Anastasia Möller, Sylviane Verschaeve, Tine Limberg"
date: "2023-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```

## Proteome-wide Screen for RNA-dependent Proteins:

## *non-synchronized A549 cells*

Loading the data:

```{r}
#MS_Table <- read.table("C:/Users/tinel/OneDrive/Dokumente/RUPRECHT-KARLS-UNIVERSITÄT HEIDELBERG/MOBI/HOCHSCHULSEMESTER4/DATAANALYSIS/RDeeP_A549_NS.csv", header=TRUE, row.names=1, sep = ";")

#MS_Table <- read.table("C:/users/schad/Documents/Bioinfoprojekt/RDeeP_A549_NS.csv", header=TRUE, r+ow.names=1, sep = ";")

#View(MS_Table)
```
=======
---
title: "Proteome-wide Screen for RNA-dependent Proteins: non-synchronized A549 cells"
author: "Johannes Schadt, Anastasia Möller, Sylviane Verschaeve, Tine Limberg"
date: "2023-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```

## Proteome-wide Screen for RNA-dependent Proteins:

## *non-synchronized A549 cells*

Loading the data:

```{r}
MS_Table <- read.table("C:/Users/tinel/OneDrive/Dokumente/RUPRECHT-KARLS-UNIVERSITÄT HEIDELBERG/MOBI/HOCHSCHULSEMESTER4/DATAANALYSIS/RDeeP_A549_NS.csv", header=TRUE, row.names=1, sep = ";")

#View(MS_Table)
```
>>>>>>> e9c16b77e10bef107667d71d131f0b7969767e23
